---
title: "Re-BLAST"
author: "Eleanor (Ella) Crotty"
header-includes:
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
output:
  pdf_document:
    toc: TRUE
urlcolor: blue
---

```{r Package Imports, message = F, warning = F}
# Warnings and startup messages suppressed
library(tidyverse)
library(patchwork)
library(scales)
library(ggrepel)
library(readxl)
library(here)
```

```{r}
# Reference
copepodNames <- read_csv(here("PMEL-Data", "OCNMS_Copepods_Krill_copy.csv"))
copepods <- c(unique(copepodNames$Species)) # this accidentally gets two krill. ask me if i give a shit.
# Import unaltered JV data
allReads <- read_csv(here("PMEL-Data", "FishPlusCOI_Reads_copy.csv")) %>% # non-cleaned allReads
  mutate(year = year(Date_local)) %>% 
  relocate(year, .after = Date_local) %>% 
  filter(Species %in% copepods | Genus %in% c("Calocalanus", "Clausocalanus", "Paracalanus") | Class == "Hexanauplia") # this will take For Fucking Ever if i run it on all species
```

```{r}
# Extract just unique ESVs and their sequences for re-BLAST ing
ESVs <- unique(allReads$ESV)
Sequences <- unique(allReads$sequence)
Re_BLAST_in <- data.frame(ESVs, Sequences)
```

```{r}
# Make a bunch of horrid little individual BLAST files <3

for (i in 1:length(ESVs)) {
  # name
  name <- ESVs[i]
  # sequence
  sequence <- Sequences[i]
  # make it into a .fsa file
  defline <- paste(">", name, sep = "")
  fsa <- data.frame(defline)
  fsa <- rbind(fsa, sequence)
  # export with ESV name
  write.table(fsa, here("BLAST_local_2025_03_14", "Re_BLAST_in", paste(name, "fsa", sep = ".")), quote = F, row.names = F, col.names = F) # all those F files keep it from adding random bullshit
}
```

-- run BLAST script that takes the .fsa file and outputs a .csv of results --

-outfmt 

#' qseqid = query sequence ID
#' sseqid = subject sequence ID
#' pident = percentage of identical matches
#' qcovs = query coverage per subject
#' sacc = subject accession
#' staxid = subject taxonomy id
#' ssciname = subject scientific name
#' sblastname = subject blast name
#' sscinames = unique subject scientific names
#' stitle = subject title

https://omicstutorials.com/step-by-step-guide-customizing-blast-output/

https://www.biostars.org/p/76551/

```
-outfmt "6 ssciname pident"
-max_target_seqs 5 
*test1 
blastn -db nt -query Re_BLAST_in/ESV_270123.fsa  -out Re_BLAST_out/ESV_270123_results.out -max_target_seqs 5 -outfmt "6 ssciname sscinames pident" -remote
*testloop
for i in Re_BLAST_in/*.fsa
  do
  echo "${i}"
  base=$(basename $i .fsa)
  echo "Re_BLAST_out/${base}_results.out"
  done
*connorsuggest
for i in *.fsa
do 
longass database query $i &
sleep 2
done
*final
for i in Re_BLAST_in/*.fsa
  do
  base=$(basename $i .fsa)
  blastn -db nt -query "${i}"  -out "Re_BLAST_out/${base}_results.out" -max_target_seqs 5 -outfmt "6 ssciname sscinames pident" -remote
  sleep 120
  done
  
  
# need to re-do a bunch that failed - their output files are 0 B
ls -lh Re_BLAST_out | grep ' 0' > Round_2.txt
awk '{ print $9 }' Round_2.txt > Round_2_names.txt

# Check progress
ls -lh Re_BLAST_out | grep ' 0' | wc -l

Round_2_names.txt is 
ESV_217525_results.out
...continues with one per line

https://stackoverflow.com/questions/16623835/remove-a-fixed-prefix-suffix-from-a-string-in-bash
$ prefix="hell"
$ suffix="ld"
$ string="hello-world"
$ foo=${string#"$prefix"}
$ foo=${foo%"$suffix"}
$ echo "${foo}"
o-wor

for line in $(cat Round_2_names.txt); 
  do
  string="${line}"
  suffix="_results.out"
  base=${string%"$suffix"}
  echo "${base}"
  done
# successfully echoes ESV_267585... one ESV per line

# now to get the in and out
for line in $(cat Round_2_names.txt); 
  do
  string="${line}"
  suffix="_results.out"
  base=${string%"$suffix"}
  echo "${base}"
  echo Re_BLAST_in/"${base}".fsa % input
  echo "Re_BLAST_out/${base}_results.out" % output
  done

Re_BLAST_Round2.sh
for line in $(cat Round_2_names.txt); 
  do
  string="${line}"
  suffix="_results.out"
  base=${string%"$suffix"}
  echo "${base}"
  blastn -db nt -query "Re_BLAST_in/"${base}".fsa"  -out "Re_BLAST_out/${base}_results.out" -max_target_seqs 5 -outfmt "6 ssciname sscinames pident" -remote &
  sleep 120
  done
  
# new plan because i need to do it locally (too many query)

Re_BLAST_Round2.sh on VM
for line in $(cat Round_2_names.txt); 
  do
  string="${line}"
  suffix="_results.out"
  base=${string%"$suffix"}
  echo "${base}"
  blastn -db nt -query "Re_BLAST_in/"${base}".fsa"  -out "Re_BLAST_out/${base}_results.out" -max_target_seqs 5 -outfmt "6 ssciname sscinames pident" -n_threads 32 &
  sleep 300
  outfile="Re_BLAST_out/${base}_results.out"
  echo head -n 1 "${outfile}"
  done
  
test echo head -n 1 ESV_273780_results.out

nohup bash Re_BLAST_Round2.sh


# Check progress
ls -lh Re_BLAST_out | grep ' 0 Mar' | wc -l

```

```{r}
# Import BLAST results
# test #1
tab <- read.table(here("BLAST_local_2025_03_14", "Re_BLAST_out", "ESV_270123_results.out"), sep = "\t") # ok great it works big loop time
# Move anything with <95% top match to the bin
top_pct <- tab[1,3]
# Take the first result of each and make a dataframe of ESVs and species
if (top_pct < 95) {
  print("Ew")
} else {
  print(tab[1,1])
}
```

```{r}
# import all
# ESVs_temp <- c("ESV_263448", "ESV_263493", "ESV_263535", "ESV_263590", "ESV_268882", "ESV_268883", "ESV_272173") # delete this line when re-BLAST finished, and just use ESVs
ESV_loop <- c()
Re_BLAST_Species <- c()
PctMatchLoop <- c()
badMatch <- c()

# from code below: 
# ESV_231703 = Pseudocalanus mimus 91.95
# ESV_220098 = Metridia pacifica 100.000
# ESV_274183 = Metridia pacifica 99.042
# ESV_264603 = Acartia longiremis 98.689
# ESV_263949 = Acartia longiremis 98.355
# ESV_271858 = Removed for 2nd match being < 95% match


for (i in 1:length(ESVs)) {
  if (ESVs[i] == "ESV_231703") { # Manually assign, didn't get BLASTed by the script
      species <- "Pseudocalanus mimus"
      name <- ESVs[i]
      top_pctl <- 91.95 # <95, so mark for removal
      badMatch <- c(badMatch, ESVs[i])
      ESV_loop <- c(ESV_loop, name)
      PctMatchLoop <- c(PctMatchLoop, "DELETE") # mark for removal in fix
      Re_BLAST_Species <- c(Re_BLAST_Species, "DELETE")
    } else { 
      if (ESVs[i] == "ESV_220098" | ESVs[i] == "ESV_274183"){ # manually assign, 2nd-3rd match
        name <- ESVs[i]
        species <- "Metridia pacifica"
        if (ESVs[i] == "ESV_220098") {
          top_pctl <- 100.000
        } else
        top_pctl <- 99.042
    } else
      if (ESVs[i] == "ESV_264603" | ESVs[i] == "ESV_263949"){ # manually assign, 2nd-3rd match
        name <- ESVs[i]
        species <- "Acartia longiremis"
        if (ESVs[i] == "ESV_264603") {
          top_pctl <- 98.689
        } else
        top_pctl <- 98.355
    } else
      if (ESVs[i] == "ESV_271858") { # first match env sample, 2nd match < 95%
        name <- ESVs[i]
        badMatch <- c(badMatch, ESVs[i])
        ESV_loop <- c(ESV_loop, name)
        PctMatchLoop <- c(PctMatchLoop, "DELETE") # mark for removal in fix
        Re_BLAST_Species <- c(Re_BLAST_Species, "DELETE")
      } else {
      # Import all the others and extract top hit and pctmatch
      # name
      name <- ESVs[i]
      #print(name)
      # import
      filename <- paste(name, "_results.out", sep = "")
      tabl <- read.table(here("BLAST", "Re_BLAST_out", filename), sep = "\t")
      #print(filename)
      species <- tabl[1,1]
      top_pctl <- tabl[1,3]
  }
  # filter for OK % match
  if (top_pctl < 95) { # Remove and take note if < 95% match
    #print(paste("Removing", name))
    badMatch <- c(badMatch, name)
    ESV_loop <- c(ESV_loop, name)
    PctMatchLoop <- c(PctMatchLoop, "DELETE") # mark for removal in fix
    Re_BLAST_Species <- c(Re_BLAST_Species, "DELETE")
  } else { # Add to key if > 95% match
      # Add species ID and pctmatch to key
      #print(species)
      ESV_loop <- c(ESV_loop, name)
      PctMatchLoop <- c(PctMatchLoop, top_pctl)
      Re_BLAST_Species <- c(Re_BLAST_Species, species)
  }
    }
}
 
re_BLAST_IDs <- data.frame(ESV_loop, Re_BLAST_Species, PctMatchLoop)

write_csv(re_BLAST_IDs, here("eDNA_Index_Hypoxia", "Data", "re_BLAST_dictionary.csv"))

#' spot check on [1] "ESV_263448"
#' [1] "ESV_263448_results.out"
#' [1] "Pseudocalanus sp. KLMPC092-09"
#' all good, checked on BLAST website
```

```{r}
allReads_reassigned <- allReads %>% 
  full_join(re_BLAST_IDs, by = join_by(ESV == ESV_loop)) %>% 
  relocate(c(Re_BLAST_Species, PctMatchLoop), .after = Species)
```

```{r}
problem_children <- allReads_reassigned %>% 
  filter(Re_BLAST_Species != Species)

# summarize problem children
problem_children_summary1 <- problem_children %>% 
  group_by(Species, Re_BLAST_Species) %>% 
  summarize(n = n(), minPctMatch = min(PctMatchLoop))

problem_children_summary <- allReads_reassigned %>% 
  filter(Species == "Oithona similis" | Species == "Acartia longiremis" | Species == "Metridia pacifica" | Species == "Pseudocalanus acuspes" | Species == "Pseudocalanus newmani") %>% 
  group_by(ESV, Species, Re_BLAST_Species) %>% 
  summarize(n = n(), minPctMatch = min(PctMatchLoop))
# looks OK - need to reassign some P. newmani, and remove some low-match ones

env_samples <- allReads_reassigned %>% 
  filter(Re_BLAST_Species %in% c("marine zooplankton environmental sample", "invertebrate environmental sample"))

env_samp_ESVs <- unique(env_samples$ESV)
```

```{r, eval = F}
# import env samples
ESV_env_loop <- c()
Re_BLAST_Species2 <- c()
PctMatchLoop2 <- c()

for (i in 1:length(env_samp_ESVs)) {
  if (env_samp_ESVs[i] == "ESV_231703") { # Manually assign this one that didn't get BLASTed by the script
    species <- "Pseudocalanus mimus"
    top_pctl <- 91.95
  } else { # Import all the others and extract top hit and pctmatch
    # name
    name <- env_samp_ESVs[i]
    #print(name)
    # import
    filename <- paste(name, "_results.out", sep = "")
    tabl <- read.table(here("BLAST", "Re_BLAST_out", filename), sep = "\t")
    #print(filename)
    species <- tabl[4,1]
    top_pctl <- tabl[4,3]
  }
  # filter for OK % match
  if (top_pctl < 95) {
      print(paste("Removing", name))
  } else {
      # Add species ID and pctmatch to key
      #print(species)
      ESV_env_loop <- c(ESV_env_loop, name)
      PctMatchLoop2 <- c(PctMatchLoop2, top_pctl)
      Re_BLAST_Species2 <- c(Re_BLAST_Species2, species)
  }
}
 
re_BLAST_IDs_2 <- data.frame(ESV_env_loop, Re_BLAST_Species2, PctMatchLoop2)
# ESV_220098 = Metridia pacifica 100.000%
# ESV_274183 = Metridia pacifica 99.042%
# ESV_264603 = Acartia longiremis 98.689%
# ESV_263949 = Acartia longiremis 98.355%
# ESV_271858 = Removed for 2nd match being < 95% match
```

