---
title: "Thesis Goal 1 (without 2023)"
author: "Eleanor (Ella) Crotty"
header-includes:
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
output:
  pdf_document:
    toc: TRUE
urlcolor: blue
---

```{r Package Imports, message = F, warning = F}
# Warnings and startup messages suppressed
library(tidyverse)
library(patchwork)
library(scales)
library(ggrepel)
library(readxl)
library(here)
library(mgcv)
library(gratia)
library(scales)
library(ggpmisc) # to put R^2 on binomial regressions
```

Goal 1: Compare copepod eDNA index, a measure of relative abundance using eDNA detections, to dissolved oxygen data in order to assess whether hypoxia decreases copepod abundance in OCNMS.

# Goals

-   Plot eDNA index over time and oxygen
-   Make scatterplots of eDNA index vs oxygen
-   Do a GAM of eDNA index vs oxygen
-   Compare eDNA to hypoxic thresholds

# Import data

```{r}
envData <- read_csv(here("PMEL-Data", "EnvironmentalDataset1_copy.csv")) %>% 
  filter(year != 2023) %>%  # Ignoring 2023 due to gaps for now
  mutate(year = as.factor(year)) %>% 
  relocate(year, .after = date)

# Fix misidentifications
re_BLAST_IDs <- read_csv(here("eDNA_Index_Hypoxia", "Data", "re_BLAST_dictionary.csv"))
allReads1 <- read_csv(here("PMEL-Data", "FishPlusCOI_Reads_copy.csv")) %>% # non-cleaned allReads
  mutate(year = year(Date_local)) %>% 
  relocate(year, .after = Date_local) %>% 
  full_join(re_BLAST_IDs, by = join_by(ESV == ESV_loop)) %>% 
  relocate(c(Re_BLAST_Species, PctMatchLoop), .after = Species) # real fix for the misidentification

allReads <- allReads1 %>% 
  filter(Re_BLAST_Species != "DELETE" | is.na(Re_BLAST_Species)) %>% # filter out <95% matches
  mutate(Species = case_when(!is.na(Re_BLAST_Species) ~ Re_BLAST_Species, .default = Species)) %>%  # Re-identify ESVs
  select(-c(Re_BLAST_Species, PctMatchLoop)) # Remove Re_BLAST_Species and PctMatchLoop bc they are no longer needed and might mess w the code

yrReads <- allReads %>% 
  filter(year != 2023) %>%   # Ignoring 2023 due to gaps for now
  mutate(year = as.factor(year))

# Reference
copepodNames <- read_csv(here("PMEL-Data", "OCNMS_Copepods_Krill_copy.csv"))
copepods <- c(unique(copepodNames$Species))
eDNAxEnvDataSat <- read_csv(here("PMEL-Data", "eDNAxEnvDataSat_copy.csv"))

# Filter to only copepods for results
allCopepodReads <- allReads %>% 
  filter(Species %in% copepods | Genus %in% c("Calocalanus", "Clausocalanus", "Paracalanus") | Class == "Hexanauplia")
```

# Results section

```{r}
unique(allCopepodReads$Field_collection_method)
unique(allCopepodReads$Amplicon)

COI <- allReads %>% 
  filter(Amplicon == "UniCOI")

copepSpResults <- allCopepodReads %>% 
  mutate(Present = case_when(nReads == 0 ~ F, .default = T)) %>% 
  group_by(year, Family, Genus, Species) %>% 
  summarize(detections = sum(Present), totReads = sum(nReads), maxReads = max(nReads), meanReads = round(mean(nReads), 2), minReads = min(nReads))

copepResults <- allCopepodReads %>% 
  mutate(Present = case_when(nReads == 0 ~ F, .default = T)) %>% 
  group_by(SampleId) %>% 
  summarize(maxReads = max(nReads), meanReads = round(mean(nReads), 2), minReads = min(nReads), detections = sum(Present))

results <- COI %>% 
  mutate(Present = case_when(nReads == 0 ~ F, .default = T)) %>% 
  group_by(SampleId) %>% 
  summarize(maxReads = max(nReads), meanReads = round(mean(nReads), 2), minReads = min(nReads), detections = sum(Present))

print("total reads")
sum(COI$nReads)
print("copepod reads")
sum(allCopepodReads$nReads)
print("copepod reads excluding 2023")
sum((allCopepodReads %>% filter(year(Date_UTC) != 2023))$nReads)

COI %>% 
  group_by(year) %>% 
  summarize(Samples = length(unique(SampleId)))

COI %>% 
  filter(Technical_Replicate == 3) %>% 
  group_by(SampleId) %>% 
  summarize(n = n())

BLAST_check <- allCopepodReads %>% 
  group_by(Class, Order, Family, Genus, Species, ESV, sequence) %>% 
  summarize(n = n())
```


# eDNA index

## Check for samples on the same day

```{r}
# Dates with 3 or 9 samples taken (see OCNMS_eDNA_Exploration.Rmd in PMEL-Hollings-eDNA-Hypoxia-2024 to see how I got this list)
duplDates <- as.POSIXct(c("2021-08-25 09:48:00", "2021-10-08 11:15:00", "2022-06-21 13:28:00", "2022-07-19 12:58:00", "2022-08-22 10:59:00", "2022-09-20 11:00:00", "2023-06-15 14:30:00", "2023-07-26 12:54:00"))

allR_check <- allReads %>% 
  filter(Date_local %in% duplDates) %>% 
  filter(ESV == "ESV_067624")
allR_check
```

## Define PCR_reps_combine

```{r}
# First, combine PCR replicates and average
PCR_reps_combine <- function(df) {
  # print(head(df))
  
  # Separate out Sample_Name into three informational columns
  df_out <- df %>% 
    separate(Sample_Name, 
             into=c("E_no", "Cruise1", "PCR_Rep"), 
             remove=F, 
             fill = "right") %>% # sep is a regular expression which is annoying, but the default recognizes any non-alphanumeric characters so the default works here
    mutate(.)
  
  # Define which columns to group by
  id_cols1 <- colnames(df_out) # Pull out column names
  id_cols1 <- id_cols1[! id_cols1 %in% c("X", "SampleId","Sample_Name", "PCR_Rep", "pctMatch", "JV_Sample_Name", "Technical_Replicate", "nReads", "Barcode.y", "E_no", "Depth_m", "Rosette_position", "Field_Sample_Name", "Field_Notes")] # Remove SampleId and Sample_Name, as well as other things that differ by PCR replicate. Also removed E_no, Depth, and Rosette_position because they differ by multiple samples in one day (exclusive to Niskin samples). These are all the things removed from id_cols.
  print(id_cols1)
  
  df_out <- df_out %>% 
    group_by_at(id_cols1) %>% # I am using species and datetime to group, which deletes E-no but I don't think I've been using it. This should be ok.
    # dplyr::mutate(SampleId = dplyr::first(SampleId)) %>% 
    dplyr::summarize(nReads = mean(nReads)) %>% 
    relocate(nReads, .after = PI)
    
  df_out
}
```



## Calculate eDNA index for 2021-2022 only

```{r}
index1dummyYr <- yrReads %>% # Quick proportions
  dplyr::group_by(SampleId) %>% 
  mutate(Tot = sum(nReads),
      	Row.sums = nReads / Tot) %>% # calculate proportions - 0 reads/0 total = NaN, need to replace with 0 to make max() work
  relocate(c(Tot, Row.sums), .after = SampleId)

yrReads_PCRcomb <- PCR_reps_combine(yrReads) # Spot check E1325 was correct

# Second, combine by species

id_cols <- colnames(yrReads_PCRcomb) # Pull out column names
id_cols <- id_cols[! id_cols %in% c("X", "ESV", "sequence", "nReads")] # Remove ESV + nreads because those are different within species

allReads_speciesYr <- yrReads_PCRcomb %>% 
  group_by_at(id_cols) %>% # group_by_at can take a vector
  dplyr::summarize(TotalnReads = sum(nReads)) %>% # Removed , avgpctMatch = mean(pctMatch) because I had to remove pctMatch to combine PCR replicates. Sum nReads results in taking the sum of all ESVs within a species
  relocate(TotalnReads, .after = PI)

write_csv(allReads_speciesYr, here("eDNA_Index_Hypoxia", "Data", "eDNA_index_no23.csv")) # Gonna want this later

allReads_speciesYr <- allReads_speciesYr %>% 
  mutate(Species_abbrev = sub("(\\w+)(\\S)", substring(Species, 1, 1), Species)) %>% # find word + space, replace with first letter
  relocate(Species_abbrev, .after = Species) %>% 
  mutate(ID_Ella = paste(Date_UTC, Amplicon, sep = "_")) %>% # just making unique IDs until grouping by it fixes the eDNA index
  relocate(ID_Ella, .after = Species) %>% 
  mutate(ID_Ella = gsub(" ", "_", ID_Ella)) %>% 
  mutate(ID_Ella = gsub("-", "", ID_Ella)) %>%
  mutate(ID_Ella = gsub(":", "", ID_Ella))

length(unique(allReads_speciesYr$ID_Ella)) # should = 68634 because that's the length of the dataset

all_index1Yr <- allReads_speciesYr %>%
  dplyr::group_by(ID_Ella) %>% # bc sampleID & E_no had to be removed in the PCR replicate combination step
  mutate(Tot = sum(TotalnReads),
      	Row.sums = TotalnReads / Tot) %>% # calculate proportions - 0 reads/0 total = NaN, need to replace with 0 to make max() work
  relocate(c(TotalnReads, Tot, Row.sums), .after = Species) # Move it somewhere I can see the damn thing

yr_eDNA_index <- all_index1Yr %>% 
  dplyr::group_by(Species) %>%
  mutate(Row.sums = case_when(Row.sums == "NaN" ~ 0, 
                                      .default = Row.sums)) %>% # Make 0/0 = 0 and not NaN
  mutate(Colmax = max(Row.sums), Normalized.reads = Row.sums / Colmax) %>%  #transforms raw number of reads to eDNA index. Creates same divide by 0 error, so:
  mutate(Normalized.reads = case_when(Normalized.reads == "NaN" ~ 0, 
                                      .default = Normalized.reads)) %>% # Make 0/0 = 0 and not NaN
  relocate(c(Colmax, Normalized.reads), .after = Row.sums)  # Move it somewhere I can see the damn thing

# Filter to only copepods
copepod_eDNA_indexYr <- yr_eDNA_index %>% 
  filter(Species %in% copepods | Genus %in% c("Calocalanus", "Clausocalanus", "Paracalanus") | Class == "Hexanauplia") %>% 
  rename(eDNA_index = `Normalized.reads`) # to make parsing this easier

write_csv(copepod_eDNA_indexYr, here("eDNA_Index_Hypoxia", "Data", "Copepod_eDNA_index_no23.csv")) # Gonna want this later

# envData = EnvironmentalDataset1
# Based on eDNAxpO2.Rmd from summer project
# copepod_eDNA_index is still essentially a version of allReads so I can use the same code I did with allReads_Species

# Round environmental data and add DateMatch
EnvRd <- envData %>% 
  mutate(DateMatch = round_date(date, unit = "10 minutes")) # Round to the nearest 10 minutes
DetectRdYr <- copepod_eDNA_indexYr %>% 
  mutate(DateMatch = round_date(Date_UTC, unit = "10 minutes"), Date_local_hr = round_date(Date_local, unit = "hour")) # Spot check - looks good. 

eDNAindxEnvDataYr <- left_join(DetectRdYr, EnvRd, by = join_by(DateMatch)) %>% 
  relocate(date, .after = Date_UTC) %>% 
  relocate(year.x, .after = Date_UTC) %>% 
  relocate(year.y, .after = Date_UTC) %>% 
  filter(year.x != 2023) %>% 
  filter()

investigateYr <- eDNAindxEnvDataYr %>% select(Species, DateMatch, Date_UTC, Date_local_hr, source, temperature, DO, Amplicon)

system("say Data Join Without 2023 Complete")

# Export the joined data
write_csv(eDNAindxEnvDataYr, here("eDNA_Index_Hypoxia", "Data", "Copepod_eDNAindxEnvData_no23.csv"))

# Make a version without the copepods that are only detected in 2023 because those plots are unhelpful
eDNAindxEnvData_cleanYr <- eDNAindxEnvDataYr %>% 
  filter(Colmax > 0) %>% 
  mutate(Present = case_when(TotalnReads == 0 ~ F, .default = T)) %>% 
  relocate(Present, .before = TotalnReads)

# Make a split version for looping over species
dfsplit <- split(eDNAindxEnvData_cleanYr, eDNAindxEnvData_cleanYr$Species) # Split by species
  
write_csv(eDNAindxEnvData_cleanYr, here("eDNA_Index_Hypoxia", "Data", "Copepod_eDNAindxEnvData_clean_no23.csv"))
```

```{r}
# Find number of copepods
eDNAindxEnvDataYr %>% 
  group_by(Family, Genus, Species) %>% 
  summarize(Reads = sum(TotalnReads))

n_species <- eDNAindxEnvDataYr %>% 
  group_by(Family, Genus, Species) %>% 
  summarize(Reads = sum(TotalnReads)) %>% 
  filter(Reads > 0) # 17
```

# Results section pt2

2021-22
```{r}
n_no23 <- copepod_eDNA_indexYr %>% 
  group_by(Species) %>% 
  filter(eDNA_index > 0) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))

max(eDNAindxEnvData_cleanYr$temperature)
```


# Plots

## Time series

```{r}
# Explore eDNA index by species
ggplot(copepod_eDNA_indexYr, aes(x = Date_local, y = eDNA_index, color = Species)) +
  geom_point(show.legend = F, size = 1) +
  facet_wrap(facets = vars(year(Date_local)), scales = "free_x") +
  geom_line(show.legend = F) +
  theme_bw()

ggplot(eDNAindxEnvDataYr, aes(x = Date_local, y = eDNA_index, color = Species)) +
  geom_point(show.legend = F, size = 1) +
  facet_wrap(facets = vars(year(Date_local)), scales = "free_x") +
  geom_line(show.legend = F) +
  theme_bw()
```
```{r}
# testing palettes lol
x <- c(1:10)
y <- x
z <- x
test <- data.frame(x, y, z)
ggplot(test, aes(x = x, y = y, color = z)) +
  geom_point() +
  scale_colour_gradient(low = "goldenrod2", high = "darkgreen") +
  theme_bw()
```


```{r}
source(here("eDNA_Index_Hypoxia", "eDNA_EnvGraphFunction.R"))

eDNAGraph(eDNAindxEnvData_cleanYr, # ignoring 2023
          envCond = "DO", 
          envCondName = "Oxygen", 
          filepath = here("eDNA_Index_Hypoxia", "Plots", "eDNAxDO_no23"),
          ylab = "Dissolved Oxygen (mg/L)", 
          widthpx = 3000, # make it longer
          threshold = F,
          thresholdLvl = 2
          )

system("say Graphs complete")
```

## Hypoxic threshold

```{r}
eDNAindxEnvData_cleanYr <- eDNAindxEnvData_cleanYr %>% 
  mutate(Hypoxic = case_when(DO < 2 ~ T, .default = F))

ggplot(eDNAindxEnvData_cleanYr, aes(x = Date_local, y = DO, color = Hypoxic)) +
  theme_bw() +
  geom_point()
```

```{r}
ggplot(eDNAindxEnvData_cleanYr, aes(x = eDNA_index)) +
  geom_histogram(fill = "cornflowerblue") +
  theme_bw() +
  facet_wrap(facets = vars(Hypoxic), ncol = 1)
```

## DO x Temp x eDNA

```{r}
# Distinguish the zero values better in the scatterplots
# I could do this by making a custom scale....ORRRR I could just add another layer where 0 is black lol :)

for (i in 1:length(dfsplit)) { # For each species: 
    species <- dfsplit[[i]]$Species[1] # Species name 
    title <- paste(species, sep = " ", "eDNA Index vs Dissolved Oxygen & Temperature") # Plot title
    print(title)
    
# Plot + save
    
    print( # plot this species
      ggplot(dfsplit[[i]], aes(x = temperature, 
                                    y = DO, 
                                    size = eDNA_index, color = eDNA_index)) +
        scale_color_viridis_c(option = "viridis", direction = -1) + # color
        scale_size(range = c(2, 7)) +
        geom_point(alpha = 0.5) +
        geom_point(shape = 1) + # add outline
        geom_point(data = dfsplit[[i]] %>% filter(eDNA_index == 0), 
                   aes(x = temperature, y = DO),
                   color = "black", size = 2, shape = 1, stroke = 1.5) + # add more distinctive zeroes
        labs(x = "Temperature (C)", 
             y = "Dissolved Oxygen (mg/L)", 
             color = "eDNA Index", 
             size = "eDNA Index",
             title = title) +
        geom_hline(yintercept = 0.66, linetype = 2, 
                   color = "red", linewidth = 1) + # most die
        geom_hline(yintercept = 0.9, linetype = 2, 
                   color = "orange", linewidth = 1) + # ~50% die
        geom_hline(yintercept = 2.66, linetype = 2, 
                   color = "forestgreen", linewidth = 1) + # sublethal effects
        theme_bw() +
        theme(text = element_text(size = 25), 
              axis.text = element_text(size = 30, color = "black"), 
              plot.title = element_text(size = 17),
              legend.title = element_text(size = 20),
              legend.text = element_text(size = 20))
    )
      
    ggsave(filename = paste(species, sep = "_", "eDNA_DO_Temp_ScatterSize.png"), 
           path = here("eDNA_Index_Hypoxia", "Plots", "DO_Temp_Scatter"), 
           width = 2500, height = 2000, units = "px")
}
```

```{r}
# Add a gray background layer for all the environmental data

for (i in 1:length(dfsplit)) { # For each species: 
    species <- dfsplit[[i]]$Species[1] # Species name 
    title <- paste(species, sep = " ", "eDNA Index vs Dissolved Oxygen & Temperature") # Plot title
    print(title)
    
# Plot + save
    
    print( # plot this species
      ggplot(data = envData, aes(x = temperature, y = DO)) +
        geom_point(color = "gray") +
        
        geom_point(data = dfsplit[[i]], aes(x = temperature, 
                                    y = DO, 
                                    size = eDNA_index, color = eDNA_index), 
                   alpha = 0.5) +
        scale_color_viridis_c(option = "viridis", direction = -1) + # color
        scale_size(range = c(2, 7)) +
        geom_point(data = dfsplit[[i]], aes(x = temperature, 
                                    y = DO, 
                                    size = eDNA_index, color = eDNA_index), 
                   shape = 1) + # add outline
        geom_point(data = dfsplit[[i]] %>% filter(eDNA_index == 0), 
                   aes(x = temperature, y = DO),
                   color = "black", size = 2) + # add more distinctive zeroes
        
        labs(x = "Temperature (C)", 
             y = "Dissolved Oxygen (mg/L)", 
             color = "eDNA Index", 
             size = "eDNA Index",
             title = title) +
        geom_hline(yintercept = 0.66, linetype = 2, 
                   color = "red", linewidth = 1) + # most die
        geom_hline(yintercept = 0.9, linetype = 2, 
                   color = "orange", linewidth = 1) + # ~50% die
        geom_hline(yintercept = 2.66, linetype = 2, 
                   color = "forestgreen", linewidth = 1) + # sublethal effects
        theme_bw() +
        theme(text = element_text(size = 25), 
              axis.text = element_text(size = 30, color = "black"), 
              plot.title = element_text(size = 17),
              legend.title = element_text(size = 20),
              legend.text = element_text(size = 20))
    )
      
    ggsave(filename = paste(species, sep = "_", "eDNA_DO_Temp_ScatterSize_AllBackground.png"), 
           path = here("eDNA_Index_Hypoxia", "Plots", "DO_Temp_Scatter"), 
           width = 2500, height = 2000, units = "px")
}
```

```{r}
for (i in 1:length(dfsplit)) { # For each species: 
    species <- dfsplit[[i]]$Species[1] # Species name 
    title <- paste(species, sep = " ", "eDNA Index vs Dissolved Oxygen & Temperature") # Plot title
    print(title)
    
# Plot + save
    
    print( # plot this species
      ggplot(dfsplit[[i]], aes(x = temperature, 
                                    y = DO, 
                                    color = eDNA_index)) +
        scale_color_continuous(type = "viridis") +
        geom_point(alpha = 0.7) +
        geom_point(data = dfsplit[[i]] %>% filter(eDNA_index == 0), 
                   aes(x = temperature, y = DO),
                   color = "black") + # Not a fan of this in the non-sized one, bc it can conceal small detections at similar DO & temp values
        ggtitle(title) +
        geom_hline(yintercept = 0.66, linetype = 2, color = "red") + # most copepods die
        geom_hline(yintercept = 0.9, linetype = 2, color = "orange") + # ~50% of copepods die
        geom_hline(yintercept = 2.66, linetype = 2, color = "forestgreen") + # many copepods experience sublethal effects (e.g. less egg production)
        theme_bw()
    )
      
    ggsave(filename = paste(species, sep = "_", "eDNA_DO_Temp_Scatter.png"), 
           path = here("eDNA_Index_Hypoxia", "Plots", "DO_Temp_Scatter"), 
           width = 2500, height = 2000, units = "px")
}
```

```{r}
for (i in 1:length(dfsplit)) { # For each species: 
    species <- dfsplit[[i]]$Species[1] # Species name 
    title <- paste(species, sep = " ", "eDNA Index vs Dissolved Oxygen & Temperature") # Plot title
    print(title)
    
# Plot + save
    
    print( # plot this species
      ggplot(dfsplit[[i]], aes(x = temperature, 
                                    y = DO, 
                                    color = eDNA_index)) +
        scale_color_continuous(type = "viridis") +
        geom_point(alpha = 0.5, shape = 15, size = 3) +
        ggtitle(title) +
        geom_hline(yintercept = 0.66, linetype = 2, color = "red") + # most copepods die
        geom_hline(yintercept = 0.9, linetype = 2, color = "orange") + # ~50% of copepods die
        geom_hline(yintercept = 2.66, linetype = 2, color = "forestgreen") + # many copepods experience sublethal effects (e.g. less egg production)
        theme_bw()
    )
      
    ggsave(filename = paste(species, sep = "_", "eDNA_DO_Temp_Heat.png"), 
           path = here("eDNA_Index_Hypoxia", "Plots", "DO_Temp_Heat"), 
           width = 2500, height = 2000, units = "px")
}
```

## DO vs. eDNA scatter plot

```{r}
# Distinguish the zero values better in the scatterplots
# I could do this by making a custom scale....ORRRR I could just add another layer where 0 is black lol :)

for (i in 1:length(dfsplit)) { # For each species: 
    species <- dfsplit[[i]]$Species[1] # Species name 
    title <- paste(species, sep = " ", "eDNA Index vs Dissolved Oxygen") # Plot title
    print(title)
    
# Plot + save
    
    print( # plot this species
      ggplot(dfsplit[[i]], aes(x = DO, y = eDNA_index)) +
        geom_point(size = 2) +
        labs(x = "Dissolved Oxygen (mg/L)", 
             y = "eDNA Index", 
             title = title) +
        geom_vline(xintercept = 0.66, linetype = 2, 
                   color = "red", linewidth = 1) + # most die
        geom_vline(xintercept = 0.9, linetype = 2, 
                   color = "orange", linewidth = 1) + # ~50% die
        geom_vline(xintercept = 2.66, linetype = 2, 
                   color = "forestgreen", linewidth = 1) + # sublethal effects
        theme_bw() +
        theme(text = element_text(size = 25), 
              axis.text = element_text(size = 30, color = "black"), 
              plot.title = element_text(size = 17),
              legend.title = element_text(size = 20),
              legend.text = element_text(size = 20))
    )
      
    ggsave(filename = paste(species, sep = "_", "eDNA_DO_Scatter.png"), 
           path = here("eDNA_Index_Hypoxia", "Plots", "eDNA_DO_Scatter"), 
           width = 2500, height = 2000, units = "px")
}
```


## Density plot

Try making a stat_density_2d(geom - "polygon") ggplot. Segment out into grids - bin the DO and temp data 

```{r}
for (i in 1:length(dfsplit)) { # For each species: 
    species <- dfsplit[[i]]$Species[1] # Species name 
    title <- paste(species, sep = " ", "eDNA Index vs Dissolved Oxygen") # Plot title
    print(title)
    
    # If no or only 1 detections, don't do it (1 detection didn't get filtered out, but it won't density)
    if (sum(dfsplit[[i]]$Present == T) <= 1) {
      print(paste(species, "not detected"))
      next
    } else
    
# Plot + save
    
    
    print( # plot this species
      ggplot(dfsplit[[i]], aes(x = DO, y = temperature)) +
        geom_point(color = "gray") + # all sampling dates
        geom_density_2d_filled(data = dfsplit[[i]] %>% 
                                 filter(Present == T),
                        aes(x = DO, y = temperature), 
                        alpha = 0.5) + # colored contours
        geom_density_2d(data = dfsplit[[i]] %>% filter(Present == T),
                        aes(x = DO, y = temperature), color = "black") + # contour lines
        geom_point(data = dfsplit[[i]] %>% filter(Present == T),
                   aes(x = DO, y = temperature), color = "black") + # Presences
        theme_bw() +
        labs(x = "Dissolved Oxygen (mg/L)", 
             y = "Temperature (C)", 
             title = title) +
        geom_vline(xintercept = 0.66, linetype = 2, 
                   color = "red", linewidth = 1) + # most die
        geom_vline(xintercept = 0.9, linetype = 2, 
                   color = "orange", linewidth = 1) + # ~50% die
        geom_vline(xintercept = 2.66, linetype = 2, 
                   color = "forestgreen", linewidth = 1) + # sublethal effects
        theme_bw() +
        theme(text = element_text(size = 25), 
              axis.text = element_text(size = 30, color = "black"), 
              plot.title = element_text(size = 17),
              legend.title = element_text(size = 20),
              legend.text = element_text(size = 20))
    )
      
    ggsave(filename = paste(species, sep = "_", "eDNA_DO_Temp_Density.png"), 
           path = here("eDNA_Index_Hypoxia", "Plots", "eDNA_DO_Temp_Density"), 
           width = 2500, height = 2000, units = "px")
}
```

## Heatmap


### Testing

```{r}
# Goal: Make a dataset with DO (binned), temp (binned), and presences/samples in that bin

testAL <- dfsplit[[1]]

testAL <- testAL %>% 
  relocate(DO, .after = Present) %>% 
  relocate(temperature, .after = DO) %>% 
  mutate(DO_bin = as.factor(round(DO*4) / 4)) %>% # note: these bins are CENTERED around a 0.5, so the actual range of the 1.0 bin is 0.75-1.25
  relocate(DO_bin, .after = DO) %>% 
  mutate(temperature_bin = as.factor(round(temperature*4) / 4)) %>% 
  relocate(temperature_bin, .after = temperature)

testAL_props <- testAL %>% 
  group_by(DO_bin, temperature_bin) %>% 
  dplyr::summarize(Samples = n(), Detections = sum(Present), Detected = Detections/Samples)

# Now plot time

ggplot(testAL_props, aes(x = DO_bin, y = temperature_bin, fill = Detected)) +
  geom_tile() +
  theme_bw()
```

### Exporting

```{r}
for (i in 1:length(dfsplit)) { # For each species: 
  
    species <- dfsplit[[i]]$Species[1] # Species name 
    title <- paste(species, sep = " ", "eDNA Detection Density") # Plot title
    print(title)
    
# Plot + save
    
    df <- dfsplit[[i]]

    df <- df %>% 
      drop_na(DO, temperature) %>% 
      relocate(DO, .after = Present) %>% 
      relocate(temperature, .after = DO) %>% 
      mutate(DO_bin = as.factor(round(DO*4) / 4)) %>% # note: these bins are CENTERED around a 0.5, so the actual range of the 1.0 bin is 0.75-1.25
      relocate(DO_bin, .after = DO) %>% 
      mutate(temperature_bin = as.factor(round(temperature*4) / 4)) %>% 
      relocate(temperature_bin, .after = temperature) 
    
    df_props <- df %>% 
      group_by(DO_bin, temperature_bin) %>% 
      dplyr::summarize(Samples = n(), Detections = sum(Present), Detected = Detections/Samples)
    
    # Now plot time
    
    print(ggplot(df_props, aes(x = DO_bin, y = temperature_bin, fill = Detected)) +
      geom_tile() +
      scale_fill_gradient(low = "skyblue", high = "darkblue") +
      ggtitle(title) +
      theme_bw())
      
    ggsave(filename = paste(species, sep = "_", "eDNA_Env_Density.png"), 
           path = here("eDNA_Index_Hypoxia", "Plots", "eDNA_Env_Density"), 
           width = 2500, height = 2000, units = "px")
}
```

# Binomial regressions

```{r}
for (i in 1:length(dfsplit)) {
  df <- dfsplit[[i]] %>% mutate(PresentBin = case_when(Present == T ~ 1, Present == F ~ 0)) %>%  # Select one species and mutate to 0/1 so that the regression will run
    filter(temperature < 11) # Filter out the outlier
  
  species <- as.character(dfsplit[[i]][1,7]) # Extract species name for title + export filename

  ggplot(df, aes(x = DO, y = PresentBin, color = PresentBin)) +
    geom_point(size = 4) +
    scale_color_gradient(high = "dodgerblue3", low = "orange2", guide = "legend", na.value = "black") + # Color to match presence/absence graphs
    geom_smooth(method = "glm", method.args = list(family = "binomial"), se = F) +
    # Print the binomial regression
    stat_poly_eq(use_label(c("R2", "P", "AIC")), size = 6, label.y = 0.75, label.x = 60) + # Print R^2 and p-value from the model
    theme_bw() +
    theme(text = element_text(size = 20)) +
    labs(title = species, x = "Dissolved Oxygen (mg/L)", y = "Species Present?")
  
    ggsave(filename = here("eDNA_Index_Hypoxia", "Plots",
                           "Binomial_Regressions", 
                           paste(species, sep = "", ".png")), # Save graph
           width = 2300, 
           height = 2000, 
           units = "px")
}
```


```{r}
system("say Script complete")
```

