---
title: "Thesis ZOIB Troubleshooting"
author: "Eleanor (Ella) Crotty"
header-includes:
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
output:
  pdf_document:
    toc: TRUE
urlcolor: blue
---

```{r Package Imports, message = F, warning = F}
# Warnings and startup messages suppressed
library(tidyverse)
library(patchwork)
library(scales)
library(ggrepel)
library(readxl)
library(here)
library(mgcv)
library(gratia)
library(ggpmisc)
library(zoib)
library(coda)

# from andrew heiss
library(brms)             # Bayesian modeling through Stan
library(tidybayes)        # Manipulate Stan objects in a tidy way
library(broom)            # Convert model objects to data frames
library(broom.mixed)      # Convert brms model objects to data frames
library(betareg)          # Run beta regression models
library(extraDistr)       # Use extra distributions like dprop()
library(ggdist)           # Special geoms for posterior distributions
library(gghalves)         # Special half geoms
library(ggbeeswarm)       # Special distribution-shaped point jittering
library(patchwork)        # Combine ggplot objects
library(marginaleffects)  # Calculate marginal effects for regression models
library(modelsummary)     # Create side-by-side regression tables
```

# Goals 

- Focus on one species first
- Make working beta GAM
- Make a zero-inflated model

```{r}
data <- read_csv(here("eDNA_Index_Hypoxia", "Data", "Copepod_eDNAindxEnvData_clean_no23.csv")) %>% # from ThesisGoal1.Rmd
  filter(!is.na(DO)) %>% 
  filter(Species == "Acartia longiremis") # you can change this to Oithona similis to check another species

dataF <- data %>% filter(temperature < 10) # no outlier
```

# ZOIB

Bayesian Inference for Beta Regression and Zero-or-One Inflated
Beta Regression

Convergence of MCMC chains can be checked using functions traceplot(), autocorr.plot() and gelman.diag() available in package coda on the posterior draws from the MCMC algorithms. zoib also contains a function check.psrf() that checks whether the multivariate psrf value can be calculated for multi-dimensional variables, provides box plots and summary statistics on multiple univariate psrf values. Posterior summary of the parameters can be obtained using summary().

When there are multiple responses y, each y should be separated by "|", such as "y1 | y2 | y3" on the left hand side of the formula. On the right side of the formula, it can include up to 5 parts in the following order: xb | xd | x0 | x1 | z, where xb represents the fixed-effects covariates/factors in the link function of the mean of the beta distribution, xd represents the fixed-effects covariates/factors in the link function of the sum of the two shape parameters of the beta distribution, x0 represents the fixed-effect covariates/factors in the link function of Pr(y=0), x1 represents the fixed-effect covariates/factors in the link function of Pr(y=1|y>0), and z represents the random-effect covariates/factors. The current version of the package only accomodates z being the same across all the link functions that have a random component. xb and xd should always be specified, even if xd contains only an intercept. If there is no zero inflation in any of the y’s, then the x0 part can be omitted, and zoib automatically adjusts the order of the rest of the X specifications; similarly for the x1 part and the random effects part z.

x1 and x2 should always be specified, even if x2 contains only an intercept (represented by 1).

For example, if there are 3 response variables y1,y2,y3 and 2 independent variables (xx1, xx2), and none of the y’s has zero inflation, then model y1 | y2 | y3 ∼ xx1 + xx2 | 1 | xx1 | xx2 implies x1 = (1,xx1,xx2), x2= 1 (intercept), x3 = NULL, x4 = (1,xx1), z = (1,xx2).

```{r}
# should I smooth it? not sure
# picked logit because it is default

# Josie Meeting
# MCMC chains start with random numbers, see if they fit the data, and then adjust back and forth until it converges
# Need the fourth thing to be |1, because there's a constant probability of one. With that, don't need one.inflation = F or setting inits.

b <- matrix(data = 1, nrow = 2*length(data$DO), ncol = 2*length(data$DO))

xd = rep(1, length(data$eDNA_index))
ind_DO_GAM_zoib <- zoib(eDNA_index ~ DO | 1 | DO | 1, # 1 must be used as an intercept if there's only one x variable, and the second 1 means there's a constant probability of a 1, not dependent on DO 
                        # using x variable as x0 because it's zero-inflated
                        data = data#, 
                        #one.inflation = F,
                        #n.burn = 1000, # burning in period of MCMC chains - the first chunk of non-converged guesses, which we will ignore. default 200.
                        #inits = list(b=NULL, # 2*n = 2*length(data$DO)?
                        #             d=NULL, # can't figure out how do do the evil matrix thing
                        #             b0=0.5, # starting value for P(y=0), or probability of non-detection 
                        #             b1=1/80, # starting value for P(y=1)
                        #             sigma=c(0.1), # sigma is a vector containing the standard deviation of each random component
                        #             R=NULL)
)
                        
# The notations b, d, b0, b1 are the same as in the "random" section above, with each specified in a matrix format of dimension (nx+1)*n.response, where nx is number of regression coefficients corresponding to the respective x’s (+1 because of the intercept). sigma is a vector containing the standard deviation of each random component, and R contains the lower triangle of the correlation matrix. For example, in a 3x3 correlation matrix, R is specified as c(1,r21,1,r31,r32,1). Each inner list contains the starting values for one MCMC chain. If initial values are specified only for a subset of the parameters in a model, NULL should be used for the rest of the unspecified parameters (whose initial values will be generated automatically by the function)

# getting invalid parent values - allegedly this is always because somehow the model is running into a value of zero, and this can be fixed by specifying inits better, but also nobody has a straight answer for how to specify inits better

coeff_zoib <- ind_DO_GAM_zoib$coeff
summary(coeff_zoib)

plot(ind_DO_GAM_zoib$coeff) # shows the MCMC chain converging
```

[1] "* List of parameter for which the posterior samples are generated         *"
[1] "* b: regression coeff in the linear predictor for the mean of beta dist'n *"
[1] "* d: regression coeff in the linear predictor for the sum of the two      *"
[1] "*    shape parameters in the beta distribution                            *"
[1] "* b0: regression coeff in the linear predictor for Prob(y=0)              *"
[1] "* b1: regression coeff in the linear predictor for Prob(y=1)              *"

[1] "Coefficients are presented in the order of b, b0 (if zero.inflation=TRUE),"
[1] "b1 (if one.inflation=TRUE), and d. If the names of independent variables X"
[1] "are not shown for the coefficients within each type (b, b0, b1, d), the   "
[1] "first coeffient is always the intercept, followed the coefficients for the"
[1] "X's in the order as how they are entered in the model specification.      "

1. Empirical mean and standard deviation for each variable,
   plus standard error of the mean:

         Mean     SD Naive SE Time-series SE
b[1]  -1.7695 0.3072 0.004435       0.004531
b[2]   0.2839 0.1110 0.001603       0.001603
b0[1] -0.8213 0.9497 0.013708       0.014554
b0[2] -0.8467 0.5087 0.007343       0.010133
b1    -4.6457 1.3093 0.018898       0.024215
d      0.8383 0.1766 0.002549       0.002625

# Meeting 4/4/2025

## From brms models, but might be helpful

https://www.andrewheiss.com/blog/2021/11/08/beta-regression-guide/#zero-inflated-beta-regression-bayesian-style

- running tidy(model) should summarize the terms
- zi_intercept is the zero-inflating component
- predicted_draws(newdata = tibble(DO seq)) should give me the predictions I want

https://mjskay.github.io/tidybayes/articles/tidy-brms.html


```{r}
x_zip <- data$DO
y_zip <- data$eDNA_index

newdata2 <- tibble(DO = c(seq(0, max(x_zip), 0.1)))

broom::tidy(ind_DO_GAM_zoib)

tidy_pred <- ind_DO_GAM_zoib %>% 
  predicted_draws(newdata = newdata2)
tidy_pred
```


```{r, eval = F}
# mgcv plotting function, shift = intercept does not need to be done, so this is probably better
species <- "Acartia longiremis"
title <- paste(species, sep = " ", "eDNA Index vs Dissolved Oxygen")
x_zip <- data$DO
y_zip <- data$eDNA_index
png(filename = paste(here("eDNA_Index_Hypoxia", "Plots", "ZOIB_Pred_test"), sep = "/", 
                         paste(species, sep = "_", "eDNA_DO_GAM_ZOIB.png"))) # open png
    
xnew1 <- data.frame(DO = c(seq(0, max(x_zip), 0.1)))
    
prediction <- pred.zoib(ind_DO_GAM_zoib, xnew = xnew1)
    
predlist <- prediction$pred
pred1 <- data.frame(predlist[[1]]) # this is a matrix with the number of rows = xnew and number of columns = Sample size per chain
# so i think i want the last column, because that's the end of the chain
predplot <- pred1[,2300]
predDO <- xnew1$DO
    
#xnew1 <- as.vector(xnew1)
predplot_df <- data.frame(predDO, predplot)
    
ggplot(predplot_df, aes(x = predDO, y = predplot)) +
    geom_point()


    
#plot <- plot(ind_DO_GAM_zoib$ypred, col = "blue") # plot GAM
# xlab = "Dissolved Oxygen (mg/L)", ylab = "eDNA Index (GAM in Blue)", main = title, 
#points(x_zip, y_zip, col = "orange", pch = 16) # plot actual data
#abline(v = 0.66, lty = "dotted", lwd = 2, col = "red") # most copepods die
#abline(v = 0.9, lty = "dotted", lwd = 2, col = "orange") # ~50% of copepods die
#abline(v = 2.66, lty = "dotted", lwd = 2, col = "forestgreen") # many copepods die
#plot
#print(plot)
#dev.off() # close png
```

